---
title: "Product Cost Change"
author: "Rising Eagle Construction LLC"
date: "June 2023"
output:
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
library(ezids)
knitr::opts_chunk$set(echo = FALSE)
```

###### Product Percentage change from 2020 to 2021



```{r include=FALSE}
library(readr)
library(ggplot2)

df <- read_csv('/Users/nusratprithee/Documents/Rising-Eagle-company-work/Purchase-History-with-Average-Unit-Price-2020-vs.-2021.csv')



```


##### Product type

```{r echo = FALSE}
# Get unique values from the "Type" column
unique_types <- unique(df$Type)

# Print each unique type
for (type in unique_types) {
  print(paste("Product Type:", as.character(type)))
}

```

#### Plot of percentage change of price

```{r echo = FALSE}


# Convert the average unit price columns from strings to numeric values
df$`Average Unit Price 01/01/21 to 12/31/21` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/21 to 12/31/21`))
df$`Average Unit Price 01/01/20 to 12/31/20` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/20 to 12/31/20`))

# Calculate the percentage change
df$`Percentage Change` <- ((df$`Average Unit Price 01/01/21 to 12/31/21` - df$`Average Unit Price 01/01/20 to 12/31/20`) / df$`Average Unit Price 01/01/20 to 12/31/20`) * 100

library(plotly)

# Create an interactive bar chart with labels and percentage change values
plot_ly(df, x = ~`Item #`, y = ~`Percentage Change`, type = "bar", text = ~paste("Type: ", `Type`), hoverinfo = "text",
        marker = list(color = "blue")) %>%
  add_text(x = ~`Item #`, y = ~`Percentage Change`, text = ~paste(sprintf("%.1f%%", `Percentage Change`)), textposition = "outside", 
           hoverinfo = "none", showlegend = FALSE) %>%
  layout(xaxis = list(title = "Item #"), yaxis = list(title = "Percentage Change"),
         title = "Percentage Change of Average Unit Price (2020 to 2021)", 
         hoverlabel = list(bgcolor = "white", font = list(color = "black")))



```

### Percentage change in Each Catagory of Product in between 2020 to 2021

```{r echo = FALSE}


# Calculate the overall percentage change for each type
overall_percentage_change <- aggregate(`Percentage Change` ~ Type, data = df, FUN = function(x) sum(x, na.rm = TRUE))

# Print the overall percentage change for each type
print(overall_percentage_change)



```

```{r echo = FALSE}
library(ggplot2)

# Calculate the overall percentage change for each type
overall_percentage_change <- aggregate(`Percentage Change` ~ Type, data = df, FUN = function(x) sum(x, na.rm = TRUE))

# Create a bar plot of the overall percentage change with labels
ggplot(overall_percentage_change, aes(x = reorder(Type, `Percentage Change`), y = `Percentage Change`)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_text(aes(label = paste0(round(`Percentage Change`, 2), "%")), vjust = -0.5) +
  labs(x = "Type", y = "Overall Percentage Change",
       title = "Overall Percentage Change of Each Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
##  Product Buy Comparison of 2020 to 2021



```{r echo = FALSE}


library(ggplot2)

# Group the data by Item #, Description, and Product Type and sum the quantity bought for 2020-2021
gpca_table <- aggregate(cbind(`Qty Sold 01/01/21 to 12/31/21`, `Qty Sold 01/01/20 to 12/31/20`) ~ `Item #` + `Description` + `Type`, data = df, sum)

# Calculate the threshold values for identifying large and medium spikes
max_qty <- max(gpca_table$`Qty Sold 01/01/21 to 12/31/21`)
large_threshold <- max_qty * 1.0  # Adjust the threshold as needed
medium_threshold <- max_qty * 0.1  # Adjust the threshold as needed

# Create a line plot of quantity bought
ggplot(gpca_table, aes(x = `Item #`, group = 1)) +
  geom_line(aes(y = `Qty Sold 01/01/21 to 12/31/21`, color = "Qty Sold 2021"), size = 1) +
  geom_point(aes(y = `Qty Sold 01/01/21 to 12/31/21`, color = "Qty Sold 2021"), size = 2) +
  geom_line(aes(y = `Qty Sold 01/01/20 to 12/31/20`, color = "Qty Sold 2020"), size = 1) +
  geom_point(aes(y = `Qty Sold 01/01/20 to 12/31/20`, color = "Qty Sold 2020"), size = 2) +
  geom_line(aes(y = `Qty Sold 01/01/21 to 12/31/21`, linetype = "White Year"), color = "black", size = 1, linetype = "dashed") +
  labs(x = "Item #", y = "Quantity Buy", title = "Quantity Buy Comparison") +
  scale_color_manual(values = c("blue", "red"), labels = c("Qty Sold 2021", "Qty Sold 2020")) +
  scale_shape_manual(values = c(16, 16)) +
  scale_linetype_manual(values = "dashed", labels = "White Year") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = ifelse(`Qty Sold 01/01/21 to 12/31/21` >= large_threshold | `Qty Sold 01/01/21 to 12/31/21` >= medium_threshold, as.character(`Type`), ""), y = `Qty Sold 01/01/21 to 12/31/21`), vjust = -0.5)




```

# Overall Cost change of product between 2020 to 2021

```{r echo = FALSE}
library(readr)



# Convert the average unit price columns from strings to numeric values
df$`Average Unit Price 01/01/21 to 12/31/21` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/21 to 12/31/21`))
df$`Average Unit Price 01/01/20 to 12/31/20` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/20 to 12/31/20`))

# Calculate the overall percentage change
overall_percentage_change <- ((sum(df$`Average Unit Price 01/01/21 to 12/31/21`) - sum(df$`Average Unit Price 01/01/20 to 12/31/20`)) / sum(df$`Average Unit Price 01/01/20 to 12/31/20`)) * 100

# Print the overall percentage change
cat("Overall Percentage Change: ", overall_percentage_change, "%", "\n")


```

```{r echo = FALSE}

# Create a data frame for the overall percentage change
data <- data.frame(Year = c("2020", "2021"), Percentage_Change = c(0, overall_percentage_change))

# Create a line graph
ggplot(data, aes(x = Year, y = Percentage_Change, group = 1)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 3) +
  labs(x = "Year", y = "Percentage Change", title = "Overall Percentage Change of Average Unit Price") +
  geom_text(aes(label = paste0(Percentage_Change, "%")), vjust = -0.5, size = 4, hjust = -0.2) +
  theme_minimal()

```

###### Product Percentage change from 2021 to 2022


```{r include=FALSE}
library(readr)
library(ggplot2)

df <- read_csv('/Users/nusratprithee/Documents/Rising-Eagle-company-work/Purchase History with Average Unit Price - 2021 vs. 2022.csv')

```

##### Product type

```{r echo = FALSE}
# Get unique values from the "Type" column
unique_types <- unique(df$Type)

# Print each unique type
for (type in unique_types) {
  print(paste("Product Type:", as.character(type)))
}

```

#### Plot of percentage change of price
```{r echo = FALSE}


# Convert the average unit price columns from strings to numeric values
df$`Average Unit Price 01/01/22 to 12/31/22` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/22 to 12/31/22`))
df$`Average Unit Price 01/01/21 to 12/31/21` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/21 to 12/31/21`))

# Calculate the percentage change
df$`Percentage Change` <- ((df$`Average Unit Price 01/01/22 to 12/31/22` - df$`Average Unit Price 01/01/21 to 12/31/21`) / df$`Average Unit Price 01/01/21 to 12/31/21`) * 100

library(plotly)

# Create an interactive bar chart with labels and percentage change values
plot_ly(df, x = ~`Item #`, y = ~`Percentage Change`, type = "bar", text = ~paste("Type: ", `Type`), hoverinfo = "text",
        marker = list(color = "blue")) %>%
  add_text(x = ~`Item #`, y = ~`Percentage Change`, text = ~paste(sprintf("%.1f%%", `Percentage Change`)), textposition = "outside", 
           hoverinfo = "none", showlegend = FALSE) %>%
  layout(xaxis = list(title = "Item #"), yaxis = list(title = "Percentage Change"),
         title = "Percentage Change of Average Unit Price (2020 to 2021)", 
         hoverlabel = list(bgcolor = "white", font = list(color = "black")))


```
### Percentage change of each catagory of product in between 2021 to 2022
```{r echo = FALSE}


# Calculate the overall percentage change for each type
overall_percentage_change <- aggregate(`Percentage Change` ~ Type, data = df, FUN = function(x) sum(x, na.rm = TRUE))

# Print the overall percentage change for each type
print(overall_percentage_change)



```

```{r echo = FALSE}
library(ggplot2)

# Calculate the overall percentage change for each type
overall_percentage_change <- aggregate(`Percentage Change` ~ Type, data = df, FUN = function(x) sum(x, na.rm = TRUE))

# Create a bar plot of the overall percentage change with labels
ggplot(overall_percentage_change, aes(x = reorder(Type, `Percentage Change`), y = `Percentage Change`)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_text(aes(label = paste0(round(`Percentage Change`, 2), "%")), vjust = -0.5) +
  labs(x = "Type", y = "Overall Percentage Change",
       title = "Overall Percentage Change of Each Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


##  Product Buy Comparison of 2021 to 2022


```{r echo = FALSE}
# Group the data by Item # and Description and sum the quantity sold for 2021-2022
gpca_table <- aggregate(cbind(`Qty Sold 01/01/22 to 12/31/22`, `Qty Sold 01/01/21 to 12/31/21`) ~ `Item #` + `Description` + `Type`, data = df, sum)

# Calculate the threshold values for identifying large and medium spikes
max_qty <- max(gpca_table$`Qty Sold 01/01/22 to 12/31/22`)
large_threshold <- max_qty * 1.0  # Adjust the threshold as needed
medium_threshold <- max_qty * 0.1  # Adjust the threshold as needed

# Create a line plot of quantity sold
ggplot(gpca_table, aes(x = `Item #`, group = 1)) +
  geom_line(aes(y = `Qty Sold 01/01/22 to 12/31/22`, color = "Qty Sold 2022"), size = 1) +
  geom_point(aes(y = `Qty Sold 01/01/22 to 12/31/22`, color = "Qty Sold 2022"), size = 2) +
  geom_line(aes(y = `Qty Sold 01/01/21 to 12/31/21`, color = "Qty Sold 2021"), size = 1) +
  geom_point(aes(y = `Qty Sold 01/01/21 to 12/31/21`, color = "Qty Sold 2021"), size = 2) +
  geom_line(aes(y = `Qty Sold 01/01/22 to 12/31/22`, linetype = "White Year"), color = "black", size = 1, linetype = "dashed") +
 labs(x = "Item #", y = "Quantity Buy", title = "Quantity Buy Comparison") +
  scale_color_manual(values = c("blue", "red"), labels = c("Qty Sold 2022", "Qty Sold 2021")) +
  scale_shape_manual(values = c(16, 16)) +
  scale_linetype_manual(values = "dashed", labels = "White Year") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = ifelse(`Qty Sold 01/01/22 to 12/31/22` >= large_threshold | `Qty Sold 01/01/22 to 12/31/22` >= medium_threshold, as.character(`Type`), ""), y = `Qty Sold 01/01/22 to 12/31/22`), vjust = -0.5)




```


# Overall Cost change of product between 2021 to 2022

```{r echo = FALSE}
library(readr)



# Convert the average unit price columns from strings to numeric values
df$`Average Unit Price 01/01/22 to 12/31/22` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/22 to 12/31/22`))
df$`Average Unit Price 01/01/21 to 12/31/21` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/21 to 12/31/21`))

# Calculate the overall percentage change
overall_percentage_change <- ((sum(df$`Average Unit Price 01/01/22 to 12/31/22`) - sum(df$`Average Unit Price 01/01/21 to 12/31/21`)) / sum(df$`Average Unit Price 01/01/21 to 12/31/21`)) * 100

# Print the overall percentage change
cat("Overall Percentage Change: ", overall_percentage_change, "%", "\n")

```

```{r}
# Create a data frame for the overall percentage change
data <- data.frame(Year = c("2021", "2022"), Percentage_Change = c(0, overall_percentage_change))

# Create a line graph
ggplot(data, aes(x = Year, y = Percentage_Change, group = 1)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 3) +
  labs(x = "Year", y = "Percentage Change", title = "Overall Percentage Change of Average Unit Price") +
  geom_text(aes(label = paste0(Percentage_Change, "%")), vjust = -0.5, size = 4, hjust = -0.2) +
  theme_minimal()

```
### Product Price Distribution Analysis

#Price Distribution indicates the product price rage we use.


### Price Distribution of 2021

```{r echo = FALSE}

# Create a histogram of the average unit prices
ggplot(df, aes(x = `Average Unit Price 01/01/21 to 12/31/21`)) +
  geom_histogram(fill = "blue", color = "black", bins = 20) +
  labs(x = "Average Unit Price", y = "Frequency", title = "Distribution of Average Unit Prices 2021") +
  theme_minimal()


```

## Price Distribution of 2022

```{r echo = FALSE}
library(ggplot2)



 ggplot(df, aes(x = `Average Unit Price 01/01/22 to 12/31/22`)) +
  geom_histogram(fill = "blue", color = "black", bins = 20) +
  labs(x = "Average Unit Price", y = "Frequency", title = "Distribution of Average Unit Prices (2022)") +
  theme_minimal()


```

# Price Distribution of 2020

```{r echo = FALSE}
library(ggplot2)

df <- read_csv('/Users/nusratprithee/Documents/Rising-Eagle-company-work/Purchase-History-with-Average-Unit-Price-2020-vs.-2021.csv')

df$`Average Unit Price 01/01/20 to 12/31/20` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/20 to 12/31/20`))

# Create a histogram of the average unit prices
ggplot(df, aes(x = `Average Unit Price 01/01/20 to 12/31/20`)) +
 geom_histogram(fill = "blue", color = "black", bins = 20) +
  labs(x = "Average Unit Price", y = "Frequency", title = "Distribution of Average Unit Prices (2020)") +
 theme_minimal()

```




