---
title: "Product Cost Change"
author: "Rising Eagle Construction LLC"
date: "June 2023"
output:
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
library(ezids)
knitr::opts_chunk$set(echo = TRUE)
```

### Product Percentage change from 2020 to 2021



```{r include=FALSE}
library(readr)
library(ggplot2)

df <- read_csv('/Users/nusratprithee/Documents/Rising-Eagle-company-work/Purchase-History-with-Average-Unit-Price-2020-vs.-2021.csv')

```
## Product type

```{r}
# Get unique values from the "Type" column
unique_types <- unique(df$Type)

# Print each unique type
for (type in unique_types) {
  print(paste("Product Type:", as.character(type)))
}

```

# Plot of percentage change of price

```{r}


# Convert the average unit price columns from strings to numeric values
df$`Average Unit Price 01/01/21 to 12/31/21` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/21 to 12/31/21`))
df$`Average Unit Price 01/01/20 to 12/31/20` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/20 to 12/31/20`))

# Calculate the percentage change
df$`Percentage Change` <- ((df$`Average Unit Price 01/01/21 to 12/31/21` - df$`Average Unit Price 01/01/20 to 12/31/20`) / df$`Average Unit Price 01/01/20 to 12/31/20`) * 100

# Create a bar chart
ggplot(df, aes(x = `Item #`, y = `Percentage Change`)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Item #", y = "Percentage Change", title = "Percentage Change of Average Unit Price (2020 to 2021)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = sprintf("%.1f%%", `Percentage Change`)), vjust = -0.5, size = 3) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(1, 1, 1, 1, "cm"))

```



```


### Product Percentage change from 2021 to 2022


```{r include=FALSE}
library(readr)
library(ggplot2)

df <- read_csv('/Users/nusratprithee/Documents/Rising-Eagle-company-work/Purchase History with Average Unit Price - 2021 vs. 2022.csv')

```

## Product type

```{r}
# Get unique values from the "Type" column
unique_types <- unique(df$Type)

# Print each unique type
for (type in unique_types) {
  print(paste("Product Type:", as.character(type)))
}

```

# Plot of percentage change of price
```{r}
columns_to_drop <- c("Type", "Qty Sold 01/01/22 to 12/31/22", "Sales $$ 01/01/22 to 12/31/22", "Qty Sold 01/01/21 to 12/31/21", "Sales $$ 01/01/21 to 12/31/21")
df <- df[, !(names(df) %in% columns_to_drop)]

# Convert the average unit price columns from strings to numeric values
df$`Average Unit Price 01/01/22 to 12/31/22` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/22 to 12/31/22`))
df$`Average Unit Price 01/01/21 to 12/31/21` <- as.numeric(gsub("[$,]", "", df$`Average Unit Price 01/01/21 to 12/31/21`))

# Calculate the percentage change
df$`Percentage Change` <- ((df$`Average Unit Price 01/01/22 to 12/31/22` - df$`Average Unit Price 01/01/21 to 12/31/21`) / df$`Average Unit Price 01/01/21 to 12/31/21`) * 100

# Create a bar chart
ggplot(df, aes(x = `Item #`, y = `Percentage Change`)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Item #", y = "Percentage Change", title = "Percentage Change of Average Unit Price (2021 to 2022)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = sprintf("%.1f%%", `Percentage Change`)), vjust = -0.5, size = 3) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(1, 1, 1, 1, "cm"))



```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
